---
title: "Genotype report"
author: "Elena Pazhenkova"
date: "2026-03-16"
output: html_document
params:
  expected_read_number: 10000000
  run_stats: "/Users/elena/PycharmProjects/ngs_pipelines/STReamGT/DIVJA240/reports/DIVJA240_reads_summary.csv"
  alleles: "/Users/elena/PycharmProjects/ngs_pipelines/STReamGT/DIVJA240/results/DIVJA240_genotypes.txt"
  positions: "/Users/elena/PycharmProjects/ngs_pipelines/STReamGT/DIVJA240/results/DIVJA240_positions.txt"

---


#Rscript -e "rmarkdown::render('example.Rmd',params=list(args = myarg))"
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
```


```{r read_arg, include=FALSE}
#Parse the parameters ans read the files
run_stats <- read.csv(params$run_stats)
libary_name <- run_stats$kit_id

alleles <- read.csv(params$alleles, sep = "\t")
positions <- read.csv(params$positions, sep = "\t")
```




input.tsv kit_id	sample_path	tags	tags_path	primers_path	fastq1_path	fastq2_path ADD: expected read number 
ngsfilter_errors: /Users/elena/PycharmProjects/ngs_pipelines/STReamGT/DIVJA240/intermediate/ngsfilter_stat.txt 
```{r}
run_stats_long <- run_stats %>%
  pivot_longer(
    cols = starts_with("reads"),
    names_to = "step",
    values_to = "reads"
  ) %>% mutate(
    step = factor(step,levels = c(
        "reads_sequenced",
        "reads_paired_filtered",
        "reads_pass_ngsfilter"), labels = c(
        "Produced",
        "Aligned",
        "Assinged to samples")))



ggplot(run_stats_long, aes(x = step, y = reads, fill = step)) +
  geom_col(width = 0.7) +
  geom_hline(
    yintercept = params$expected_read_number,
    linetype = "dashed",
    size = 1
  ) + 
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = paste("Read counts for kit", libary_name),
    x = "",
    y = "Number of reads"
  ) +
  theme_bw(base_size = 14) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) 
```
* as produced reads we count only one-direction (F or R) reads. Dashed line shows number of reads ordered in sequencing company.


```{r}
# chunk to check ngsfilter errors.
```



number  of reads and alleles per locus/per replicate

```{r}
reads_per_marker <- alleles %>% group_by(Marker) %>% summarise(sum_reads = sum(Read_Count))
```

```{r}
sequences_per_marker <- alleles %>% group_by(Marker) %>% summarise(n_seq = n_distinct(Sequence))
```


```{r}
df <-  alleles
rows <- LETTERS[1:8]
df$row <- rows[(df$Position - 1) %/% 12 + 1]
df$col <- (df$Position - 1) %% 12 + 1


agg <- df %>%
  filter(called == TRUE) %>%
  group_by(Plate, Marker, row, col) %>%
  summarise(
    reads = sum(Read_Count),
    .groups = "drop"
  )
```



```{r}
# select plate and marker
d <- agg %>% filter(Plate == 1, Marker == "UA_03")

fig <- plot_ly(
  d,
  x = ~col,
  y = ~row,
  z = ~reads,
  type = "heatmap",
  colorscale = "Viridis",
  showscale = TRUE
) %>%
  add_trace(
    x = d$col,
    y = d$row,
    type = "scatter",
    mode = "text",
    text = d$reads,
    textposition = "middle center",
    textfont = list(color = "white"),
    hoverinfo = "none",
    showlegend = FALSE
  ) %>%
  layout(
    title = "Plate 1 — UA_03",
    xaxis = list(side = "top", dtick = 1, title = list(text = "")),
    yaxis = list(scaleanchor = "x", categoryorder = "array", categoryarray = rev(LETTERS[1:8]), title = list(text = ""))
  )

fig
```

```{r}
library(plotly)
library(dplyr)

plot_plate <- function(data, plate_id, marker_id){
  d <- data %>% filter(Plate == plate_id, Marker == marker_id)
  
  fig <- plot_ly(
    d,
    x = ~col,
    y = ~row,
    z = ~reads,
    type = "heatmap",
    colorscale = "Viridis",
    showscale = TRUE,
    colorbar = list(title = "") # remove legend title
  ) %>%
    add_trace(
      x = d$col,
      y = d$row,
      type = "scatter",
      mode = "text",
      text = d$reads,
      textposition = "middle center",
      textfont = list(color = "white"),
      hoverinfo = "none",
      showlegend = FALSE
    ) %>%
    layout(
      title = paste(plate_id),
      xaxis = list(side = "top", dtick = 1, title = list(text = "")),
      yaxis = list(scaleanchor = "x",
                   categoryorder = "array",
                   categoryarray = rev(LETTERS[1:8]),
                   title = list(text = ""))
    )
  
  fig
}
```

```{r}
plates <- sort(unique(agg$Plate))
marker_id <- "UA_03"

fig_list <- lapply(plates, function(p){
  plot_plate(agg, plate_id = p, marker_id = marker_id)
})
```

```{r}
fig_facet <- subplot(
  fig_list,
  nrows = 2,
  margin = 0.05,
  shareX = TRUE,
  shareY = TRUE,
  titleX = TRUE,
  titleY = TRUE
)

fig_facet
```




```{r}
plates <- sort(unique(agg$Plate))
markers <- sort(unique(agg$Marker))
```


```{r}
plot_plate_marker <- function(data, plate_id, marker_id){
  d <- data %>% filter(Plate == plate_id, Marker == marker_id)
  
  plot_ly(
    d,
    x = ~col,
    y = ~row,
    z = ~reads,
    type = "heatmap",
    colorscale = "Viridis",
    showscale = TRUE,
    colorbar = list(title = ""),
    visible = FALSE,           # all traces invisible by default
    name = paste0("Plate ", plate_id)
  ) %>%
    add_trace(
      x = d$col,
      y = d$row,
      type = "scatter",
      mode = "text",
      text = d$reads,
      textposition = "middle center",
      textfont = list(color = "white"),
      hoverinfo = "none",
      showlegend = FALSE,
      visible = FALSE
    )
}
```



```{r}
n_traces <- length(markers) * length(plates)
menu_buttons <- list()

for(i in seq_along(markers)){
  visible_vector <- rep(FALSE, n_traces)
  start <- (i-1)*length(plates) + 1
  end <- i*length(plates)
  visible_vector[start:end] <- TRUE
  
  menu_buttons[[i]] <- list(
    method = "update",
    args = list(list(visible = visible_vector)),
    label = markers[i]
  )
}

fig <- fig %>%
  layout(
    updatemenus = list(
      list(
        active = 0,
        buttons = menu_buttons,
        x = 0.1,
        y = 1.15,
        xanchor = "left",
        yanchor = "top"
      )
    )
  )
```


```{r}
# set first marker visible
fig$x$data[1:(length(plates)*2)] <- lapply(fig$x$data[1:(length(plates)*2)], function(tr){
  tr$visible <- TRUE
  tr
})
```


```{r}
fig
```

